paper1_figures

### 2023-10-12 paper1_markdown.Rmd
12.10

During last years, the interest in alternative ways of social organization augmented drastically because of the social and ecological crises we are facing [@Euler2018]. Commons are one of this possible solutions, since they represent an alternative way of organizing resources who does not involve the classic private and public model [@Ostrom1990]. In her seminal work, Ostrom refers to commons (or "common-pool resource") as a resource system equally shared between a self-managed community of users, because of the difficulty of excluding somebody from benefiting from its use [@Ostrom1990]. Beyond the resource aspect of the commons, the notion of social practices (referred as commoning) is also essential [@Euler2018]. Commoning is the process through which people manage to find solutions for their needs by self-organizing independently from the state or the market. Community is therefore engaged in defining rules to successfully manage a shared resource [@Meroni2022], and that is why taking into account the specific community who is commoning is also essential to fully understand a commons [@Dove2019]. @Bollier2019 are even more radical: for them, the term "commons" should not be defined as a resource at all, but as "living social systems" who allow people to self-organize to address shared problems. Finally, "commons can thus be understood as collaborative arrangements for value production processes that are based on participatory principles" [@Seravalli2018a, p. 1].

@@

Cooperation has also extensively been studied through social dilemma. In such situations, each individual obtains a higher personal payoff with a non-cooperative, self-interested behaviour than with a social, cooperative one, independently from what others do. However, everybody's payoff is higher if all cooperate than if all defect [@Dawes1980]. This approach is very useful to gain insights into how cooperation works because it provides a standard experimental paradigm to test under which conditions people cooperate [@Balliet2011].

While economic and evolutionary theories say that individuals are always pursuing their self-interest (either by defecting to maximize their personal payoff, either by cooperating to increase their social status and reputation), social dilemma experiments' results show that it is not always the case. First, it is observed that a significant amount of people are "strong reciprocators": they are willing to use their personal resources to reward cooperators and punish defectors, even if they are not certain that their altruistic behaviour will be rewarded [@Fehr2002; @Gachter2009]. Moreover, according to the Greed Efficiency Fairness hypothesis, people's greed to pursuit their personal interest is constrained by their desire to use the resource efficiently and to fairly allocate it between members [@Wilke1991]. These results shows that cooperation is not always driven by personal interest and that people care about others' and environment's welfare [@Torsvik2011; @VanVugt2009].

@@

Given the exploratory nature of this paper, a scoping review approach is used [@Munn2018]. To conduct it, different scientific publications were analyzed in response to the research questions. To gather the papers, a Boolean search methodology was used in different scientific search engines such as Scopus, Web of Science and ScienceDirect. The key phrases were "social dilemma", "public good", "common resource", "commons", "cooperation", "conflicts", "reciprocity", "trust", and they were paired with "AND" and "OR" Boolean operators. The research was made during the month of August, 2023. According with @Munn2022, the type of source was not a selection criteria, therefore journal articles, conference papers and book chapter could be included. This allows us to have an idea as complete as possible about the existing literature on cooperation in social dilemmas and commons. Moreover, there were not any chronological or geographical limitations, once again to provide a broader vision of the topic. There were however some selection criteria concerning the papers' methodology and content. First, as mentioned in the introduction, papers investigating how an external entity/institution shapes cooperation were not included. Secondly, papers that studied social dilemmas with an agent-based model methodology were also excluded. Moreover, in case of experimental or correlative studies, only cases with adult participants were selected.

@@

node [shape = rectangle]       
rec1 [label = 'Papers identified from search engines
n = 5210', pos='0,0!', width = 4, height = 0.5]
rec2 [label = 'Duplicates
n = 624', pos='4,0!', width = 2, height = 0.5]
rec3 [label = 'Papers screened based on titles and abstracts
n = 4586', pos='0,-1!', width = 4, height = 0.5]
rec4 [label = 'Papers not suitable
n = 4333', pos='4,-1!', width = 2, height = 0.5]
rec5 [label = 'Papers read completely
n = 253', pos='0,-2!', width = 4, height = 0.5]
rec6 [label = 'Papers not suitable
n = 163', pos='4,-2!', width = 2, height = 0.5]
rec7 [label = 'Studies included
n = 90', pos='0,-3!', width = 4, height = 0.5]

@@

The papers' topics can be divided in 4 categories: first, "social dilemma setting" (*N*=76) refers to high controlled laboratory experiments where participants have to take cooperative decision. Usually some variables are manipulated to investigate the influence of a specific factor on cooperation [e.g. @Zhang2019]. The type of paradigm used was most of the time a public good game (where participants can decide how much to contribute to a public good, e.g. @Torsvik2011) or a common resource game (where participants can decide how much to take from a common resource, e.g. @Nockur2020). Moreover, the type of game could vary during the experiment [e.g. @Fosgaard2019]. The second category is "real-life social dilemma situation" (*N*=7), and investigate cooperation in different real-life situations that have a social dilemma structure such as recycling behaviours [@Rompf2017] or use of public transportation [@VanLange1998]. Thirdly, "real-life commons" (*N*=4) refers to studies that investigate cooperation with participants that are members of a commons in their everyday life. Finally, papers in the category "commons modeled by a social dilemma experiment" (*N*=3) use a laboratory social dilemma paradigm, but with clear references to a concrete common. For instance, in @Mosler1993, participants had to imagine to be fishermen who shared a common basin.

@@

Finally, most of the studies were conducted in Europe (*N*=39) (especially in Germany (*N*=12) and in the Netherlands (*N*=9)), followed by North America (*N*=20, of which *N*=19 in the USA) and Asia (*N*=9, of which *N*=6 in China). Moreover, *N*=7 studies were set in more than one country, and *N*=13 were meta-analysis or literature reviews. The two remaining studies took place in Australia and in Colombia, respectively.

@@

The partners' gender also have an impact on women's and men's cooperation. In general, groups composed only by men are more cooperative than only-women groups. This is explained by the fact that women are less achievement oriented and more socially oriented than men: they are therefore less concerned by the maximization of their personal payoff and more about relative payoff. Thus, the defection strategy allows them to play defensively and to avoid earning less than the other players [@Colman2018]. Moreover, men displayed higher levels of gratitude and trust in same-sex groups than women [@Peshkovskaya2019]. On the other side, women cooperate more in mixed group. In fact, when both men and women are present, gender stereotypes are activated and people tend to adapt their behaviour to them. This leads women to behave in a more prosocial, altruistic way and men to be more self-oriented and competitive [@Balliet2011]. The same result is also found by @Peshkovskaya2019, who also demonstrated that women reported a stronger social connection with their partners in mixed than in same-sex groups. The overall group composition also have an impact on cooperation: in general, groups with a majority of women display the highest levels of cooperation [@Spadaro2023]. However, @Barrero-Amortegui2021 showed that the level of resource exploitation is linked to the number of women in a group, but that this link is mediated by the signature of an agreement about extraction. More precisely, in their experiment, groups with a majority of women extracted more resources from a common good (and thus cooperated less) before the signature of the agreement, but also showed the largest reduction of the exploitation level after the agreement was signed. Therefore, even the effect of group composition seems to be subjected to moderator variables.

@@

## Group identity

@@

Personal values and personality traits can play an important role in predicting cooperative behaviour [@Hilbig2018]. One of the most studied is the Social Value Orientation (SVO), that divides people in pro-social and pro-self based on how much they value their own and others' outcomes in situations of interdependence [@Balliet2009]. The effect of SVO on cooperation is very consensual: in a meta-analysis, 81 out of the 82 papers examined reported than pro-social cooperated significantly more that pro-self [@Balliet2009].

@@

Through a scoping review approach, the current study identified some factors that could impact cooperation in social dilemmas and in commons. Some of the this elements can be related to Ostrom's principles to successfully manage a commons [@Ostrom1990]. In fact, her first statement says that commons need clearly defined boundaries: therefore, it must be clear who has the right to benefit from the common resource and who does not. This could be linked to the social identity construct, since only when an in- and an out-group exist social identity is salient and therefore an identification is possible [@Tajfel1974]. Even if Ostrom developed this principle on a more organizational level, the secondary effect of creating a social identity for commons' members can also contribute to improve cooperation. Moreover, individuals who are part of the group, and who are therefore affected by the commons' rules, must be able to participate in modifying them (principle 3). This statement shows the importance of social norms, and especially of injunctive ones, namely what is morally approved in a social group [@Cialdini1990]. Moreover, in order to modify such norms, people have to communicate [@Bicchieri2002]. Communication is also essential for principle 6, who posits that members of a commons must have access to low-cost arenas to resolve conflicts. In fact, communication helps to reach an agreement to the strategy that needs to be followed [@Hopthrow2005]. Finally, Ostrom's principles 4 and 5 underline the importance of incentives, especially in the form of a monitoring system implemented by members them-selves and a graduate sanctions system toward violators of the rules. The monitoring system refers to social incentives, and is linked to anonymity, since if people know that their behaviour can be observed, they are more willing to conform to social norms because they are worried about others' approval [@Nelissen2013; @Rege2004]. On the other hand, a sanction system can also have a positive impact on cooperation [@Balliet2011a], especially if it is graduate, so if sanctions are not too high at the first violation [@Rockenbach2019].[^3]

@@

Ostrom's assumptions are therefore confirmed, especially by factors at the organizational level. This is understandable when considering her economic approach to the study of the commons. However, taking into account the social dynamics and the individuals who interact within the commons is also essential to have a better understanding of how these communities work [@Euler2018; @Nightingale2011]. For this reason, this paper makes a step further by identifying some individual factors that affect the efficiency of commons.

@Tajfel1974: classic TIS, explanation of why more group identity = more cooperation, see also experiments with minimal groups paradigm + @Fritsche2018: The collectively interchangeable state of "social identity" makes people susceptible to adopt collective appraisals, goals, and behavior tendencies that are shared within their ingroup as their own

At a certain point say that these factors are not independent, but interconnected =\> prepare a model/framework

@Ackermann2019: link between SVO and trust

@@

@Chou2013: link between trust and collective efficacy, so that trust could also fit in SIMPEA and SIMCA

bystander effect =\> anonymity

@Aksoy2019: see if class or group had a stronger impact =\> could be a reference to say that it is important to create a common social identity that makes weaker the effect of previous belongings **???**

@Hogg2006: communication and social identity

@Navarro-Carillo2018: high status more generalized trust, low status more in-group depersonalized trust (because more social networks)

individual factor: gender is interpretable by social status and ss is interpretable by group identity

social status: low cooperate more than high =\> theories of Iacoviello etc!!!\
@Vargas-Salfate2018: more system justification in high-status individuals!

group factors: communication linked to social norms and social identity + anonymity underlies all the others + incentives has dark sides

@@

## Limitation

observation: majority of papers included are about SD, and very few studies about commons are quantitative --\> future research: quantitative approach to investigate cooperation in common initiatives

critiques at SD because it is artificial, does not take into account all real-life factors --\> so it is good to keep it as a base, but investigations in commons are needed

60/91 papers in Europe and North America + reviews also based on studies in these countries =\> bias (talk about cultural differences in cooperation)

>>>>

During last years, the interest in alternative ways of social organization augmented drastically because of the social and ecological crises we are facing [@Euler2018]. Commons are one of this possible solutions, since they represent an alternative way of organizing resources who does not involve the classic private and public model [@Ostrom1990]. In her seminal work, Ostrom refers to commons (or "common-pool resource") as a resource system equally shared between a self-managed community of users, also because of the difficulty of excluding somebody from benefiting from its use [@Ostrom1990]. Beyond the material aspect of the commons, the notion of social practices (referred as commoning) is also essential [@Euler2018]. Commoning is the process through which people manage to find solutions for their needs by self-organizing independently from the state or the market. Community is therefore engaged in defining rules to successfully manage a shared resource [@Meroni2022], and that is why taking into account the specific community who is commoning is also essential to fully understand a commons [@Dove2019]. @Bollier2019 are even more radical: for them, the term "commons" should not be defined as a resource at all, but as "living social systems" who allow people to self-organize to address shared problems. Finally, "commons can [...] be understood as collaborative arrangements for value production processes that are based on participatory principles" [@Seravalli2018a, p. 1].

@@

Cooperation has also extensively been studied through social dilemma. In such situations, each individual obtains a higher personal payoff with a non-cooperative, self-interested behaviour than with a social, cooperative one, independently from what others do. However, everybody's payoff is higher if all cooperate than if all defect [@Dawes1980]. This approach is very useful to gain insights into how cooperation works because it provides a standard experimental paradigm to test under which conditions people cooperate [@Balliet2011]. While economic and evolutionary theories say that individuals are always pursuing their self-interest (either by defecting to maximize their personal payoff, either by cooperating to increase their social status and reputation), social dilemma experiments' results show that it is not always the case. First, it is observed that a significant amount of people are "strong reciprocators": they are willing to use their personal resources to reward cooperators and punish defectors, even if they are not certain that their altruistic behaviour will be rewarded [@Fehr2002; @Gachter2009]. Moreover, according to the Greed Efficiency Fairness hypothesis, people's greed to pursuit their personal interest is constrained by their desire to use the resource efficiently and to fairly allocate it between members [@Wilke1991]. These results shows that cooperation is not always driven by personal interest and that people care about others' and environment's welfare [@Torsvik2011; @VanVugt2009].

@@

Given the exploratory nature of this paper, a scoping review approach is used [@Munn2018]. To conduct it, different scientific publications were analyzed in response to the research questions. To gather the papers, a Boolean search methodology was used in different scientific search engines such as Scopus, Web of Science and ScienceDirect. The key phrases were "social dilemma", "public good", "common resource", "commons", "cooperation", "conflicts", "reciprocity", "trust", and they were paired with "AND" and "OR" Boolean operators. The research was made during the month of August, 2023. According with @Munn2022, the type of source was not a selection criteria, therefore journal articles, conference papers and book chapter could be included. This allows us to have an idea as complete as possible about the existing literature on cooperation in social dilemmas and commons. Moreover, there were not any chronological or geographical limitations, once again to provide a broader vision of the topic. There were however some selection criteria concerning the papers' methodology and content. First, as mentioned in the introduction, papers investigating how an external entity/institution shapes cooperation were not included. Secondly, human participants were needed, so papers who studied social dilemmas with an agent-based model methodology were also excluded. Moreover, in case of experimental or correlative studies, only cases with adult participants were selected.

@@

node [shape = rectangle]       
rec1 [label = 'Papers identified from search engines
n = 5210', pos='0,0!', width = 4, height = 0.5]
rec2 [label = 'Duplicates
n = 624', pos='4,0!', width = 2, height = 0.5]
rec3 [label = 'Papers screened based on titles and abstracts
n = 4586', pos='0,-1!', width = 4, height = 0.5]
rec4 [label = 'Papers not suitable
n = 4333', pos='4,-1!', width = 2, height = 0.5]
rec5 [label = 'Papers read completely
n = 253', pos='0,-2!', width = 4, height = 0.5]
rec6 [label = 'Papers not suitable
n = 163', pos='4,-2!', width = 2, height = 0.5]
rec7 [label = 'Papers included
n = 90', pos='0,-3!', width = 4, height = 0.5]

@@

The papers' topics can be divided in 4 categories: first, "social dilemma setting" (*N*=76) refers to high controlled laboratory experiments where participants had to take cooperative decision. Usually some variables were manipulated to investigate the influence of a specific factor on cooperation [e.g. @Zhang2019]. The type of paradigm used was most of the time a public good game (where participants could decide how much to contribute to a public good, e.g. @Torsvik2011) or a common resource game (where participants could decide how much to take from a common resource, e.g. @Nockur2020). Moreover, the type of game could vary during the experiment [e.g. @Fosgaard2019]. The second category is "real-life social dilemma situation" (*N*=7), and investigated cooperation in different real-life situations with a social dilemma structure such as recycling behaviours [@Rompf2017] or use of public transportation [@VanLange1998]. Thirdly, "real-life commons" (*N*=4) refers to papers who studied cooperation with participants that were members of a commons in their everyday life. Finally, papers in the category "commons modeled by a social dilemma experiment" (*N*=3) used a laboratory social dilemma paradigm, but with clear references to a concrete common. For instance, in @Mosler1993, participants had to imagine to be fishermen who shared a common basin.

@@

Finally, most of the studies were conducted in Europe (*N*=39) (especially in Germany (*N*=12) and in the Netherlands (*N*=9)), followed by North America (*N*=20, *N*=19 of which in the USA) and Asia (*N*=9, *N*=6 of which in China). Moreover, *N*=6 studies were set in more than one country, and *N*=14 were meta-analysis or literature reviews. The two remaining studies took place in Australia and in Colombia, respectively.

@@

The partners' gender also have an impact on women's and men's cooperation. In general, groups composed only by men are more cooperative than only-women groups. This is explained by the fact that, in general, women are less achievement oriented and more socially oriented than men: they are therefore less concerned by the maximization of their personal payoff and more about relative payoff. Thus, the defection strategy allows them to play defensively and to avoid earning less than the other players [@Colman2018]. Moreover, men displayed higher levels of gratitude and trust in same-sex groups than women [@Peshkovskaya2019]. On the other side, women cooperate more in mixed group. In fact, when both men and women are present, gender stereotypes are activated and people tend to adapt their behaviour to them. This leads women to behave in a more prosocial, altruistic way and men to be more self-oriented and competitive [@Balliet2011]. The same result is also found by @Peshkovskaya2019, who addiotionally demonstrated that women reported a stronger social connection with their partners in mixed than in same-sex groups. The overall group composition also have an impact on cooperation: in general, groups with a majority of women display the highest levels of cooperation [@Spadaro2023]. However, @Barrero-Amortegui2021 showed that the level of resource exploitation is linked to the number of women in a group, but that this link is mediated by the signature of an agreement about extraction. More precisely, in their experiment, groups with a majority of women extracted more resources from a common good (and thus cooperated less) before the signature of the agreement, but also showed the largest reduction of the exploitation level after the agreement was signed. Therefore, even the effect of group composition seems to be subjected to moderator variables.

@@

## Group identification

@@

Personal values and personality traits can play an important role in predicting cooperative behaviour [@Hilbig2018]. One of the most studied is the Social Value Orientation (SVO), that divides people in pro-social and pro-self based on how much they value their own and others' outcomes in situations of interdependence [@Balliet2009]. The effect of SVO on cooperation is very consensual: in a meta-analysis, 81 out of the 82 papers examined reported than pro-social cooperated significantly more than pro-self [@Balliet2009].

@@

Through a scoping review approach, the current study identified several factors that could impact cooperation in social dilemmas and in commons. Some of this elements can be related to Ostrom's principles to successfully manage a commons [@Ostrom1990]. In fact, her first statement says that commons need clearly defined boundaries: therefore, it must be clear who has the right to benefit from the common resource and who does not. This could be linked to the social identity construct, since only when an in- and an out-group exist social identity is salient and therefore an identification is possible [@Tajfel1974]. Even if Ostrom developed this principle on a more organizational level, the secondary effect of creating a social identity for commons' members can also contribute to improve cooperation. Moreover, individuals who are part of the group, and who are therefore affected by the commons' rules, must be able to participate in modifying them (principle 3). This statement shows the importance of social norms, and especially of injunctive ones, namely what is morally approved in a social group [@Cialdini1990]. Moreover, in order to modify such norms, people have to communicate [@Bicchieri2002]. Communication is also essential for principle 6, who posits that members of a commons must have access to low-cost arenas to resolve conflicts. In fact, communication helps to reach an agreement to the strategy that needs to be followed [@Hopthrow2005]. Finally, Ostrom's principles 4 and 5 underline the importance of incentives, more precisely in the form of a monitoring system implemented by members them-selves and a graduate sanctions system toward violators of the rules. The monitoring system refers to social incentives, and is linked to anonymity, since if people know that their behaviour can be observed, they are more willing to conform to social norms because they are worried about others' approval [@Nelissen2013; @Rege2004]. On the other hand, a sanction system can also have a positive impact on cooperation [@Balliet2011a], especially if it is graduate, so if sanctions are not too high at the first violation [@Rockenbach2019].[^3]

@@

Ostrom's assumptions are therefore confirmed, especially by factors at the organizational level. This is understandable when considering her economic approach to the study of the commons. However, taking into account the social dynamics and the individuals who interact within the commons is also essential to have a better understanding of how these communities work [@Euler2018; @Nightingale2011]. For this reason, this paper makes a step further by identifying some individual characteristics affecting the efficiency of commons. The effects of these factors seem to be coherent with the existing literature.

For instance, the effect of trust on cooperation

-   link of every individual factor with the existing literature

    -   trust: being trusted increase the level of oxytocin, and this impact in turn the cooperative behaviour toward the person who trusted first @Campbell2010 + dispositional trust predicted pro social behaviour + mediated the relation between early-life experience and pro social behaviour @Wu2020 + conditional reciprocity = cooperate just if others do so (strong reciprocity, @Fehr2022, @VanLange1999): if no information on others, we must base our choice in trust

    -   SVO: pro-socials cooperate more because they accord more importance at others' outcome + joint outcome (vs pro-selves = competitors and individualist) @VanLange1999, integrative model of social value orientation (by according importance at the role of equality in outcomes for pro-social, this model allows to explain why pro-social can behave even less cooperatively than pro-self when reciprocity is violated)

    -   @Tajfel1974: classic TIS, explanation of why more group identity = more cooperation, see also experiments with minimal groups paradigm + @Fritsche2018: The collectively interchangeable state of "social identity" makes people susceptible to adopt collective appraisals, goals, and behavior tendencies that are shared within their ingroup as their own

    -   social status: low cooperate more than high =\> theories of Iacoviello etc!!!\
        @Vargas-Salfate2018: more system justification in high-status individuals!

    -   @Janssen2015: confirmation of several factors (to read!!!)

-   group factors: communication linked to social norms and social identity + anonymity underlies all the others + incentives has dark sides

-   saying that these factors are not independent, but interconnected

    -   For example, women level of cooperation is predicted by their trust in their interaction partner, while men's trust and cooperation are not linked [@Irwin2015]. Moreover, women's lower cooperation is explained by the fact that they expected their partner to contribute less to the common good [@Dorrough2019].

    -   @Ackermann2019: link between SVO and trust

    -   @Navarro-Carillo2018: high status more generalized trust, low status more in-group depersonalized trust (because more social networks) =\> link with gender and trust

    -   link between gender and social status, and between social status and social identity

    -   @Aksoy2019: see if class or group had a stronger impact =\> could be a reference to say that it is important to create a common social identity that makes weaker the effect of previous belongings **???**

    -   @Hogg2006: communication and social identity

    -   @Chou2013: link between trust and collective efficacy, so that trust could also fit in SIMPEA and SIMCA

    -   de-individualized trust = more trust in people of in-group just because they are in-group

    -   @VanLange1999 =\> women were more prevalent between pro-socials and less between pro-selves (secondary result)

-   in-group identification is also the main predictor of collective action in SIMCA and SIMPEA

@@

@@

## Limitations, future perspectives and conclusion

This review's aim was to provide an overview of the factors that affect cooperation in social dilemma situations such as commons. However, most of the studies included had a laboratory social dilemma setting, who do not allow to exactly reproduce the dynamics present in a commons [@Cole2010]. Therefore, this papers offers a good basis about elements influencing cooperation in commons, but future studies should confirm this results in real-life settings. Moreover, this review is not informative about the relative importance of each factor. To shed light on this topic, quantitative studies are needed.

Another limitation of the current review is the lack of representation of the global population. In fact, two thirds of the analyzed papers were studies conducted in Europe or in North America, and the fourteen reviews were also mostly based on studies from Global North. The current results apply therefore to Western societies and should not be generalized to Global South. In fact, there are cultural differences in the levels of individualism and collectivism: generally, North American and European cultures are more individualistic while South American, African and Asian ones are more collective [@Triandis1989]. These differences affect the extent to which individuals define them-selves as part of a group [@Trafimow1991] and, consequently, their behavior in cooperative situations [e.g. @Kuwabara2014]. For this reason, it is likely that underlying mechanism of cooperation are different for Global South and North: comparative studies between this two regions should be conducted in order to reach a global understanding of cooperation.

Finally, the papers included in this review did not provide a longitudinal perspective on the study of cooperation, leaving therefore some open questions: which factors are more important in the early or in the late stages of cooperation? How the different factors evolve and influence each-others over time? That is why longitudinal investigations about the evolution of commons are needed.

To conclude, this paper offers a first draft of why people do (not) cooperate in commons, and highlights a lack of systematic studies about cooperation in commons. However, it is not without limitations, thus further researches are necessary to reach a deeper understanding of the topic. + say the first important practical implication that we can already deduce (social identity, ...)
